---
title: "431 Bootstrapping Problem"
author: "Nik Floden"
date: "April 23, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1
a) 6.965201  \newline
b) 0.8516212 \newline
c) 0.0005531136 \newline
d) Normal:( 5.296,  8.634 ), Percentile:( 5.429,  8.751 )  \newline
e) The normal, the bootstrap of the ratio appears follow a normal distribution around the bootstrapped value. \newline
f) 5 is not within the 95% confidence interval, so it is appropriate to reject the null that $H_0: \mu_r = 5$ in favour of the alternate hypothesis that $\mu_r \neq 5$. \newline
g)
```{r}
set.seed(4292019)
library(boot)
x <- c(30,45,120,15,90,30,15,13,30,30,90,60,120,45)
y <- c(120,240,540,150,420,180,210,180,210,240,420,300,540,270)
n = length(x)
xy <- data.frame(cbind(x,y))
Brep = 10000
sampr <- function(xy,i=c(1:n)){
  z <- xy[i,]
  return(mean(z$y/z$x))
}
bootr <- boot(data=xy, statistic = sampr, R=Brep)
bootr
boot.ci(bootr, conf = .95)
hist(bootr$t,main = "Nik Floden")
plot(ecdf(bootr$t), main = "Nik Floden")
```

#2
a) 5.484311 \newline
b) 0.4458228 \newline
c) 0.07993244 \newline
d) Normal: ( 4.559,  6.276 ), Percentile: ( 4.905,  6.610 )
e) I would choose the percentile, the bootstrap of the ratio does not appear to follow a normal distribution. 
f) 5 is within (4.905, 6.610), so we fail to reject the null that $\mu_R = 5$ at $\alpha = 0.05$
g)
```{r}
sampbigr <- function(xy,i=c(1:n)){
  z <- xy[i,]
  return(mean(z$y)/mean(z$x))
}
bootbigr <- boot(data=xy, statistic = sampbigr, R=Brep)
bootbigr
boot.ci(bootbigr, conf = .95)
hist(bootbigr$t,main = "Nik Floden")
plot(ecdf(bootbigr$t), main = "Nik Floden")
```

#3
a)Intercept = 24.4575, Slope = 0.8923 \newline
b)Normal: ( 63.7, 137.0 ), Percentile: ( 60.0, 133.6 ) \newline
c)Normal: ( 3.124,  4.046 ), Percentile: ( 3.085,  3.995 )\newline
d) I would report the normal in (b), and percentile in (c), because the intercept follows a normal distribution and the slope does not. 
e) The hypothesized beta values are $\beta_0 = 0$ and $\beta_1 = 5$. 
f)0 is not in the confidence interval so we reject the null hypothesis that $\beta_0 = 0$. 
g)5 is not in the confidence interval, so we reject the null hyptothesis that $\beta_1 = 5$. 
h)
```{r}
#########1
lm1 <- lm(y ~ x)
summary(lm1)
#########
intercept <- function(xy, i=c(1:n)){
  z <- xy[i,]
  slr <- lm(z$y ~ z$x)
  return(slr$coef[1])
}
bootint <- boot(data=xy, statistic = intercept, R=Brep)
bootint
boot.ci(bootint, conf = .95)

par(mfrow=c(1,2))
hist(bootint$t,main = "Nik Floden")
plot(ecdf(bootint$t), main = "Nik Floden")
#########
slope <- function(xy, i=c(1:n)){
  z <- xy[i,]
  slr <- lm(z$y ~ z$x)
  return(slr$coef[2])
}
bootslope <- boot(data=xy, statistic = slope, R=Brep)
bootslope
boot.ci(bootslope, conf = .95)
par(mfrow=c(1,2))
hist(bootslope$t,main = "Nik Floden")
plot(ecdf(bootslope$t), main = "Nik Floden")
```