---
title: "STAT 431 HW4"
author: "Nik Floden"
date: "February 7, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 
A)
```{r}
rep <- 50000

Before <- c(89, 90, 87, 98, 120, 85, 97, 88, 95, 94)
After <- c(76, 101, 84, 86, 105, 84, 93, 80, 95, 90)
D <- After - Before

medianD <- median(D)
medianD

Fp = 0
upper <- 0
lower <- 0 
n = length(D)

medianDvec <- 1:rep
for(i in 1:rep){
  sgnvec <- sign(runif(n)-.5)
  Dvec <- sgnvec*D
  medianDvec[i] <- median(Dvec)
  if(medianDvec[i] >= medianD) upper = upper+1
  if(medianDvec[i] <= medianD) lower = lower+1
}

pval_lower <- lower/rep
pval_lower
```

B) With a p-value of 0.0698, there is little to no evidence that the After-Before median difference is less than zero. 

C)
```{r}
pval_upper <- upper/rep
sgnD <- sign(medianD)
if(sgnD < 0) pval_two_sided = 2*pval_lower
if(sgnD > 0) pval_two_sided = 2*pval_upper
pval_two_sided
```

D) With a p-value of .1396, there is little to no evidence to support the alternate hypothesis that the After-Before Median Difference $\neq 0$

#2
A)
```{r}
library(gtools)

ScoreA <- c(5.1, 9.0, 7.2, 8.1, 8.8, 6.6, 5.9, 6.0, 7.8, 5.5)
ScoreB <- c(5.2, 4.9, 6.1, 5.9, 7.0, 6.4, 5.7, 7.0, 4.8)

wilcox.test(ScoreA, ScoreB)
```

B) With a P-value of 0.08613, there is some evidence, but not enough evidence to reject the null that there is no difference between the two mean nesting heights.

C)
```{r}
wilcox.test(ScoreA, ScoreB, conf.int = T)
```

The true mean difference is between -0.19999 and 2.39999. 

