---
title: "STAT 431 HW 9"
author: "Nik Floden"
date: "April 10, 2019"
output: pdf_document
---

#1
a) The appropriate test is Fisher's exact test. 

b)
```{r}
x <- as.matrix(c(26,18,6,10,18,30))
dim(x) <- c(3,2)
x
fisher.test(x, alternative = "less")
```

With a p-value of .000007, we reject the null hypothesis that the true proportion of plants in good condition is equal across months. 

#2
a)
```{r}
y <- as.matrix(c(9,13,12,50))
dim(y) <- c(2,2)
y
fisher.test(y)
```

With a p-value of 0.08265 there is little evidence to reject the null that there is no difference between proportion of men hired and women hired. 

b)
```{r}
MHtable <- array(c(2,11,49,8,4,1,1,0,3,1,3,1), dim=c(2,2,3), dimnames = list(
    Applicants = c("Male", "Female"),
    Hire = c("Hired", "Not Hired"),
    College = c("Education", "Bus. Admin.", "Engineering")))
MHtable
mantelhaen.test(MHtable,correct = FALSE)
```

With a p-value of 0.07, there is suggestive, but ultimately inconclusive evidence against the null that the true proportions of men hired and women hired are equal. 

c) Yes. 

#3
a) The Kruskal-Wallis test.

b)$H_0: M_{initial} = M_1 = M_2 = M_3 = M_4$
\newline
$H_a: M_{initial} > M_1 > M_2 > M_3 > M_4$

c)
```{r}

init <- c(82, 80, 75, 65, 77, 68, 70, 77)
m1 <- c(84, 80, 78, 72, 74, 69, 74, 76)
m2 <- c(77, 76, 77, 68, 72, 65, 71, 78)
m3 <- c(76, 76, 74, 65, 75, 66, 70, 72)
m4 <- c(79, 78, 72, 66, 75, 64, 68, 70)

total <- c(init,m1,m2,m3,m4)
group <- c(rep("a",length(init)),rep("b",length(m1)), rep("c",length(m2)), rep("d",length(m3)), rep("e",length(m4)))
group <- as.factor(group)
kruskal.test(total, group)
```

With a p-value of 0.4835, there is no evidence to reject the null that the true median heartrate is equal across time goes on. 

d) No, there is no randomization of treatment that can be seen with the data provided. 

#4
```{r}
library(modeltools)
library(coin)
IN <- c(19,25,30,34,37,46,47,51,56,57,61,66,67,74,78,86,122,123,130,133,134,136,141,143,151,152,153,154,156,162,164,165,182,189)

IP <- c(22,23,38,42,73,77,89,115,144)

total <-c(IN, IP)
group <- c(rep("IN", length(IN)),rep("IP", length(IP)))
group <- as.factor(group)
median_test(total ~ group, conf.int=T, distribution="exact")
```

With a p-value of 0.4566, there is very little evidence to reject the null that the median time to death between Immunoperoxidase Negative subjects and Immunoperoxidase Positive subjects is not equal. 

#5
a)
```{r}
W <- c(37,37,33,33,30,27,27,23,23,-9,17,17,17,12,7,7,7,7,-18,-19,-1,-5,-5,-9,-9,-12,-14,-27,-17,-29,-29,-25,-25,-28,-30,-35,-36,-37,-38)
M = 18
P = 11

wobs = sum(W[1:M])
wobs
absW = abs(wobs)

iter = 100000
pW <- 1:iter

for(i in 1:iter){
  s <- sample(W,M)
  Wsum = sum(s)
  pW[i] <- Wsum
}
pW <- sort(pW)

xquantile <- quantile(pW, probs = c(.005, .01, .025, .05, .95, .975, .99, .995))

pdist <- quantile(pW, probs = seq(0,1,1/iter))

absW = abs(wobs)

ntail1 <- length(pdist[pW <= -absW])
tail1 <- ntail1/iter
tail1

ntail2 <- length(pdist[pW >= absW])
tail2 <- ntail2/iter
tail2

twotail = tail1 + tail2
twotail

plot(ecdf(pW), main="Monte Carlo ECDF of Gehan W")
```

With a p-value of 2e-05, we have sufficent evidence to reject the null that the survival between subjects using the 6-MP drug and the placebo is equal.

b) 
```{r}
library(survival)

group<- c(rep("6-MP", 18), rep("Placebo", 21))
time <- c(6, 6, 6, 7, 10, 13, 22, 23, 6, 9, 10, 17, 19, 19, 25, 32, 34, 35, 1, 1, 2, 2, 3, 4, 4, 5, 5, 8, 8, 8, 8, 11, 11, 12, 12, 15, 17, 22, 23)

censor <- c(1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1)

patient.data <- data.frame(group, time, censor)
patient.data
survdiff(Surv(time, censor == 1)~group, data = patient.data, rho = 0)
```

With a p-value of 2e-04, we reject the null that the true survival times between patients who used 6-MP and a placebo is equal. 