---
title: "Stat 431 HW6"
author: "Nik Floden"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1
The null hypothesis is the median length of the lines is the same across all samples. 
The alternate is that at least one sample has a different median length. 

\newpage



#2a
```{r}
library(perm)

prep = 50000

y <-c(2.974, 0.945, 1.639, 0.039, 1.296, 0.768, 0.803, 0.216, 0.074, 1.508, 4.825, 2.252,1.561, 2.045, 1.096)

nvec <- c(5,5,5)

group <- c(1,1,1,1,1,2,2,2,2,2,3,3,3,3,3)
group <- as.factor(group)

permControl = permControl(nmc=prep, seed = 2202019, p.conf.level = 0.95)
permKS(y, group, control=permControl)

```
#2b
The ANOVA has a p-value of 0.08834 and the Permutation test has a p-value of 0.07266. These are both suggestive, but inconclusive. More data is required to make clearer statements. 


#3
The null hypothesis is that the median phosphorus content will be the same across all heights of grass.
The alternate hypothesis is that there is at least one height of grass where the median phosphorus content less than the rest of the heights. 

```{r}
library(clinfun)

height <- c(rep(20, 7), rep(15, 7), rep(10, 7), rep(5,7))
pcon <- c(13.0, 21.1, 11.7, 16.3, 15.5, 18.5, 22.1, 24.5, 18.0, 14.0, 24.2, 11.6, 19.0, 18.5, 15.6, 23.8, 24.4, 24.0, 21.0, 21.1, 20.4, 35.3, 22.5, 16.9, 25.0, 23.1, 26.0, 22.2)

jonckheere.test(pcon, height, alternative="decreasing")
```

With a P-value of 0.001396, we have strong enough evidence to reject the null under  $\alpha = 0.5$. With this, we can state that there is at least one height of grass that has a lower phosphorus content than the rest of the heights. 